---
title: "Assignment 4 - Heart rate, respiration and interpersonal coordination"
author: "Katrine Nymann"
date: "November 20, 2019"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Heart rate, respiration and interpersonal coordination

Physiological data (here heart rate [variability], and respiration) are increasingly popular. Historically treated as pernicious noise to be regressed out of neuro-imaging data, there is now increasing research on how these signals tell us something important about cognition and beyond being just a signal of cognitive processes also impact them in interesting ways. Advanced sport science, and the quantified self movement (closely followed by marketing and communication) have hailed continuous physiological tracking as a powerful way to access and modify attitudes, habits, and performance. Further, as team coordination (in the military, in decision processes and organizational contexts) is more and more in focus, research has attempted to measure how interpersonal coordination between physiological systems might tell us something important about e.g. emotional and cognitive coordination. See references in the reading list for more on this.

In this assignment, you will learn to:
- collect physiological data
- pre-process physiological data (and grow further your mad R skills)
- model the continuous interdependence between two signals (using a multilevel model as proxy for a dynamical system approach)
- conservatively assess the presence of coordination between to signals in a controlled context

This assignment has two parts. The first part familiarizes you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from previous years (Study1, Study2 and Study 3). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: in the first year it was self-paced joint reading; in the second year it was the tv-series conversation.

## Let's get started

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal.
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3). There are also smarter packages, like cowplot and ggpubr.
- Can you eye-ball which condition if any displays more physiological coordination?

### First we read one data file and identify the procedure
- Load the file
- correctly identify all columns
- plot the data
- deal with the artifacts
- downsample the dat
- Add a column for study, group, trial and condition

```{r}
# Load the libraries
library(pacman)
pacman::p_load(ggplot2)
p_load(gridExtra, groupdata2, tidyverse)
library(stringi)
library(lme4)
# Load the file
sync <- read.csv("data/Study4_G3_T3_Synchronous.csv")
# We plot the data

#We plot respiration data for both participants
plot_resp <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, Resp1, color = "P1")) +
  geom_path(aes(sync$TimeMs, Resp2, color = "P2")) +
  labs(x = "Time", y = "Resp") +
  theme(legend.position="bottom")
plot_resp

#We plot heart rate data for both participants
plot_hr <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, HR1, color = "P1")) +
  geom_path(aes(sync$TimeMs, HR2, color = "P2")) +
  labs(x = "Time", y = "Heart Rate") +
  theme(legend.position="bottom")
plot_hr 
# We make the two plots on top of eachother
gridExtra::grid.arrange(plot_resp, plot_hr)

## Remove outliers


### Tip, check the function below
removeOuts <- function(ts,threshold){
  SD_plus=(mean(ts,na.rm=T) + sd(ts,na.rm=T)*2.5)
  SD_minus=(mean(ts,na.rm=T) - sd(ts,na.rm=T)*2.5)
  
  ts[ts > (mean(ts,na.rm=T) + (threshold*sd(ts,na.rm=T)))] = SD_plus
  ts[ts < (mean(ts,na.rm=T) - (threshold*sd(ts,na.rm=T)))] = SD_minus # This works so that values that where bigger/smaller than 2.5*SD away from the mean are now equal to SD*2.5 + mean
  return(ts)
}
threshold=2.5 # Default value at 2.5 sds from the mean
#We try with the respiration data 
sync$Resp1_r <- removeOuts(sync$Resp1, 2.5)
sync$Resp2_r <- removeOuts(sync$Resp2, 2.5)
#We do it with heart rate
sync$HR1_r <- removeOuts(sync$HR1, 2.5)
sync$HR2_r <- removeOuts(sync$HR2, 2.5)
#We plot to see how it worked
compare1 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, Resp1, color = "Original")) +
  geom_path(aes(sync$TimeMs, Resp1_r, color = "Removed")) +
  labs(x = "Time (ms)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Participant 1")

compare2 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, Resp2, color = "Original")) +
  geom_path(aes(sync$TimeMs, Resp2_r, color = "Removed")) +
  labs(x = "Time (ms)", y = "Respiration") +
  theme(legend.position="bottom")+ ggtitle("Participant 2")

compare3 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, HR1, color = "Original")) +
  geom_path(aes(sync$TimeMs, HR1_r, color = "Removed")) +
  labs(x = "Time (ms)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Participant 1")

compare4 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, HR2, color = "Original")) +
  geom_path(aes(sync$TimeMs, HR2_r, color = "Removed")) +
  labs(x = "Time (ms)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Participant 2")

# Plot raw data againt those with the artiacts removed
#We plot all visualizations together in a grid
gridExtra::grid.arrange(compare1, compare2, compare3, compare4)

## We scale the data (and overwrite to the original columns, så these are the ones we will use)
sync$Resp1 <- scale(sync$Resp1_r)
sync$Resp2 <- scale(sync$Resp2_r)
sync$HR1 <- scale(sync$HR1_r)
sync$HR2 <- scale(sync$HR2_r)

### Tip: if scale() gives some issues, try the one below
#z_scale <- function(column){
#  column_c <- (column - mean(column)) / sd(column)
#}

# Plot again to check how scaled data look like
#We plot to see how it worked
s1 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, Resp1, color = "Scaled")) +
  geom_path(aes(sync$TimeMs, Resp1_r, color = "Not-scaled")) +
  labs(x = "Time (ms)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Participant 1")

s2 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, Resp2, color = "Scaled")) +
  geom_path(aes(sync$TimeMs, Resp2_r, color = "Not scaled")) +
  labs(x = "Time (ms)", y = "Respiration") +
  theme(legend.position="bottom")+ ggtitle("Participant 2")

s3 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, HR1, color = "Scaled")) +
  geom_path(aes(sync$TimeMs, HR1_r, color = "Not scaled")) +
  labs(x = "Time (ms)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Participant 1")

s4 <- ggplot(data = sync) +
  geom_path(aes(sync$TimeMs, HR2, color = "Sacled")) +
  geom_path(aes(sync$TimeMs, HR2_r, color = "Not scaled")) +
  labs(x = "Time (ms)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Participant 2")

#We plot all visualizations together in a grid
gridExtra::grid.arrange(s1, s2, s3, s4)

## Downsample
### This is tricky, so you can have a look at my code  (relying on Ludvig's groupdata2) if you get stuck
#We make a column with each 
sync$rowname <- rep(1:177323)

p_load(tidyverse)
#We downsample
d1 = sync %>%
 group(n = 100, method = 'greedy') %>%
 dplyr::summarise(
   time = mean(TimeMs,na.rm=T),
   HR1 = mean(HR1,na.rm=T),
   HR2 = mean(HR2,na.rm=T),
   Resp1 = mean(Resp1,na.rm=T),
   Resp2 = mean(Resp2,na.rm=T),
   rowname = rowname[1]) #the index we use to put them back together

## Plot the downsampled data
d_resp <- ggplot(data = d1) +
  geom_path(aes(time, Resp1, color = "P1")) +
  geom_path(aes(time, Resp2, color = "P2")) +
  labs(x = "time (ms)", y = "Resp") +
  theme(legend.position="bottom") + ggtitle("Respiration data")

# We plot the heart rate data
d_HR <- ggplot(data = d1) +
  geom_path(aes(time, HR1, color = "P1")) +
  geom_path(aes(time, HR2, color = "P2")) +
  labs(x = "time (ms)", y = "Heart rate") +
  theme(legend.position="bottom") + ggtitle("Heart rate data")

#We plot them together
gridExtra::grid.arrange(d_HR, d_resp)

# We see that we get some flat lines. Probably from loose belts


```
## Now add the group, trial, condition to the cleaned up, scaled, downsampled data
## Tip the info is in the file name
```{r}
# We need to make a function that can read the data, remove outliers, and downsampling

#### Function for preprocessing all data 
read_filename <- function(filename){
     
    # load data
     data <- read.delim(filename, sep = ",")
     
          #In one study (study 4) there is no time column, instead it's called TimeMs (except for one       where it's time but in minutes.) Therefore we added this to rename that column for study 4
     study <- stringr::word(filename, start = 1, sep = "_") #Takes out the filename in a new          variable study
     study <- stringr::str_sub(study, start = 6) #Takes the study
     if (study=="4"){data <- data %>% rename(time = TimeMs)} #If study is study 4 it renames
    
     data$time <- as.numeric(data$time)
     
     if (study=="4"){data$time <- data$time/1000} # changing study 4 to seconds      
     #we downsample
     data$rowname<-rep(1:nrow(data),1)
     
     data = data %>%
      group(n = 1000, method = 'greedy') %>%
      dplyr::summarise(
        time = mean(time,na.rm=T),
        HR1 = mean(HR1,na.rm=T),
        HR2 = mean(HR2,na.rm=T),
        Resp1 = mean(Resp1,na.rm=T),
        Resp2 = mean(Resp2,na.rm=T),
        rowname = rowname[1])
    
     #we remove outliers
     data$Resp1<-removeOuts(data$Resp1,2.5)
     data$Resp2<-removeOuts(data$Resp2,2.5)
     data$HR1<-removeOuts(data$HR1,2.5)
     data$HR2<-removeOuts(data$HR2,2.5)
     
     #we scale
     data$Resp1<-scale(data$Resp1)
     data$Resp2<-scale(data$Resp2)
     data$HR1<-scale(data$HR1)
     data$HR2<-scale(data$HR2)
     
    # parse filename to extract study, diagnosis, subject and trial
     data$group <- word(filename, start = -3, sep = "_")
     data$group<-str_sub(data$group, start = 2)
     
     data$condition <- word(filename, start = -1, sep = "_")
     data$condition<-str_sub(data$condition, end = -5)
     
     data$study <- word(filename, start = 1, sep = "_")
     data$study<-str_sub(data$study, start = 6)
     
     data$trial <- word(filename, start = 3, sep = "_")
     data$trial <-str_sub(data$trial, start = 2)
     
     data$.groups = NULL
     data$rowname = NULL
     
     return(data)
}

#Test on one file 
test <- read_filename("Study1_G1_T1_Synchronous.csv")


#  Identify all files to be read
# Run the function on the whole dataset using map_df
listed<- list.files(pattern="*.csv") %>% 
    purrr::map_df(read_filename) 

#Creating unique ID for each pair acros studies 
listed <- listed %>% mutate(uPairID = paste(study, group, sep = "_"), 
                    uPairID = as.numeric(as.factor(uPairID)))

write.csv(listed, "data_study_4.csv")

```


## Now we are ready to go to load and pre-process all files

Go through all the files (with a function passed onto map_df), check which files should be excluded, if any, and save the pre-processed time-series

A couple of tips:
- looping is oh so slow. Making a function and using Map/Map_df is your salvation.
- each study restarts the group numbering, so you should make sure to change that (e.g. 100 * Study + Group)
- you need to make sure all the data are meaningful or something has to be removed. Plotting is your friend. E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs

```{r, include=FALSE}
# Now we need to make sure all the data are meaningful or something has to be removed
# E.g. "Study1_G1_T1_Synchronous" has one bad respiration signal. We could replace it with NAs

study_check <- read.csv("Study4_G1_T1_Synchronous.csv") ## This belt was too loose, so if we are going to analyze respiration we could replace it with NAs if we needed it. 

check_plot <- ggplot(study_check) + geom_path(aes(TimeMs, Resp1, color = "Participant 1")) +
  geom_path(aes(TimeMs, Resp2, color = "Participant 2")) +
  labs(x = "time (ms)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 1 Respiration")
check_plot
```


```{r}
#We load out df created before
phys_data <- read.csv("pre/data_study_4.csv")

# We should look especially on study 4
#For group 7
#Filter out
study_4 <- phys_data %>% filter(condition == "Synchronous" & group == "7")
#Plot
study_4_plot <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes(time, Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Synchronous")
study_4_plot

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "Synchronous" & group == "8")
#Plot
study_4_plot_2 <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes((time), Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Synchronous")
study_4_plot_2

## For the other condition

#Filter out
study_4 <- phys_data %>% filter(condition == "Conversation" & group == "7")
#Plot
study_4_plot_3 <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes(time, Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Conversation")

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "Conversation" & group == "8")
#Plot
study_4_plot_4 <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes(time, Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Conversation")

## For the third condition

#Filter out
study_4 <- phys_data %>% filter(condition == "TurnTaking" & group == "7")
#Plot
study_4_plot_5 <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes(time, Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Conversation")

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "TurnTaking" & group == "8")
#Plot
study_4_plot_6 <- ggplot(study_4) + geom_path(aes(time, Resp1, color = "Participant 1")) +
  geom_path(aes(time, Resp2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Respiration") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Conversation")

#We plot them together two and two
gridExtra::grid.arrange(study_4_plot, study_4_plot_2)
gridExtra::grid.arrange(study_4_plot_3, study_4_plot_4)
gridExtra::grid.arrange(study_4_plot_5, study_4_plot_6)

```
Now for Heart Rate data
```{r}
# We choose to only look at speaking data
# We should look especially on study 4
#For group 7
#Filter out
study_4 <- phys_data %>% filter(condition == "Synchronous" & group == "7")
#Plot
study_4_plot <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes(time, HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Synchronous")

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "Synchronous" & group == "8")
#Plot
study_4_plot_2 <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes((time), HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Synchronous")

## For the other condition

#Filter out
study_4 <- phys_data %>% filter(condition == "Conversation" & group == "7")
#Plot
study_4_plot_3 <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes(time, HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Conversation")

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "Conversation" & group == "8")
#Plot
study_4_plot_4 <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes(time, HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Conversation")

## For the third condition

#Filter out
study_4 <- phys_data %>% filter(condition == "TurnTaking" & group == "7")
#Plot
study_4_plot_5 <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes(time, HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 7 Conversation")

#For group 8
#Filter out
study_4 <- phys_data %>% filter(condition == "TurnTaking" & group == "8")
#Plot
study_4_plot_6 <- ggplot(study_4) + geom_path(aes(time, HR1, color = "Participant 1")) +
  geom_path(aes(time, HR2, color = "Participant 2")) +
  labs(x = "time (s)", y = "Heart Rate") +
  theme(legend.position="bottom") + ggtitle("Study 4 Group 8 Conversation")

#We plot them together
gridExtra::grid.arrange(study_4_plot, study_4_plot_2)
gridExtra::grid.arrange(study_4_plot_3, study_4_plot_4)
gridExtra::grid.arrange(study_4_plot_5, study_4_plot_6)

```
We see that group 7 in the synchronous condition has a flat point for heart rate for participant two in the middle.(We each did two groups each and this was the only one with visisble flat lines). We should exclude this condition for that pair. 
```{r}
# Remove bad data

#First we remove the movement data
phys_data <- phys_data %>% filter(condition == "TurnTaking" | condition == "Synchronous" | condition == "Conversation")

# We remove group 7 condition synchronous
try <- subset(phys_data, group!=7)
try_2 <- subset(phys_data, group==7 & condition!="Synchronous")
phys_data <- merge(try, try_2, all = T)

# Save the data
```



## Now we need to run some analysis

Let's start with a multilevel model that accounts for 
- stability (how each signal is autocorrelated)
- interpersonal dependence (each signal is dependent from the previous state of the other signal)

The data needs to be further prepared, so we can analyze both participants in the same model.
We need to turn the data into a long format:
- a column indicating own hr and one own respiration
- a column indicating other hr and one other respiration
- a column indicating change in hr from previous round and one in respiration

We can then run an analysis where change is a function of one's previous state (stability, see slides), and the other's previous state (coupling). Make sure to:
- set up the most interesting contrasts: how do these parameters vary by condition? which condition should be baseline?
- set up the right random effects.
- N.B. the model will be slow. Make sure it works on a subset of the data first!

Bonus question: what if we include an additional layer? Is my heart rate just adjusting to yours, or also to how much you are adjusting to mine?
- to start answering this we can add a column indicating the previous change in hr in the other and one in respiration
- we can then build on the previous models by also adding the previous change in the other


```{r}
# Genearate a column for each: previous HR1, HR2, Resp1, Resp2
## We use group by first

phys_data <- phys_data %>% group_by(condition, study, group) %>% mutate(
  HR1_past = lag(HR1, 1),
  HR2_past = lag(HR2, 1),
  Resp1_past = lag(Resp1, 1),
  Resp2_past = lag(Resp2, 1),
)
  
# Genearate a column for each: change in HR1, HR2, Resp1, Resp2
phys_data$HR1_change <- phys_data$HR1-phys_data$HR1_past
phys_data$HR2_change <- phys_data$HR2-phys_data$HR2_past
phys_data$Resp1_change <- phys_data$Resp1-phys_data$Resp1_past
phys_data$Resp2_change <- phys_data$Resp2-phys_data$Resp2_past

## Remember to predict from HR_past

# Make the data long, so we can analyze both participants at the same time 
## N.B. This is a bit tricky and you might have to do it in several steps

d_hr_change <- gather(phys_data, participant, HR_change_self, HR1_change, HR2_change)%>%
  select(time, HR_change_self, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant)) 

#We need to make an HR_change_other column
d_hr_change_other <- gather(phys_data, participant, HR_change_other, HR2_change, HR1_change)%>%
  select(time, HR_change_other, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant)) 

##### We make HR_self and HR_other as we need them later #####

HR_self<- gather(phys_data, participant, HR_self, HR1, HR2)%>%
  select(time, HR_self, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant))

HR_other <- gather(phys_data, participant, HR_other, HR2, HR1)%>%
  select(time, HR_other, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#We will make one for HR_self past 
d_hr_self <- gather(phys_data, participant, HR_self_past, HR1_past, HR2_past)%>%
  select(time, HR_self_past, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant)) 

#We will make one for HR_other past
d_hr_other <- gather(phys_data, participant, HR_other_past, HR2_past, HR1_past)%>%
  select(time, HR_other_past, participant, study, group, condition) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant)) 

#We need to add the two column to our other dfs
d_hr_change$HR_change_other <- d_hr_change_other$HR_change_other
d_hr_self$HR_other_past <- d_hr_other$HR_other_past
d_hr_self$HR_other <- HR_other$HR_other
d_hr_self$HR_self <- HR_self$HR_self

#Now we have the dataframe for d_resp_change

d_resp_change <- gather(phys_data, participant, Resp_change_self, Resp1_change, Resp2_change) %>% select(time, Resp_change_self, participant, study, group, condition) %>% mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant))

### For respiration
#adding for other
#We need to make an HR_change_other column
d_resp_change_other <- gather(phys_data, participant, Resp_change_other, Resp2_change, Resp1_change)%>%
  select(time, Resp_change_other, participant, study, group, condition, Resp1_past, Resp2_past) %>% 
  mutate(participant = parse_number(as.character(group)) * 10 + parse_number(participant)) 
d_resp_change$Resp_change_other <- d_resp_change_other$Resp_change_other

# Now we need to merge
dd <- merge(d_hr_change, d_resp_change)
dd <- merge(dd, d_hr_self)

write.csv(dd, "dd.csv")

# Model change as a function of own and other previous state 
#model <- lmer(HR_change_self ~ 1 + (HR_self + HR_other) + condition + 
  #(1 + (HR_self + HR_other) + condition | participant) + 
  #(1 + (HR_self + HR_other) + condition | group), data = dd)
#summary(model)

library(lmerTest)

# We use : to better interpret the interaction
model_c <- lmer(HR_change_self ~ 0 + condition + 
                (HR_self_past + HR_other_past):condition + 
                (0 + condition | participant) +
                (0 + condition | group), data = dd)

#Also run it with the star * so we can see whether the interaction is there

model_s <- lmer(HR_change_self ~ 0 + condition + 
                (HR_self_past + HR_other_past)*condition + 
                (0 + condition | participant) +
                (0 + condition | group), data = dd)

#I get no p values
summary(model_c)
summary(model_s)

# Bonus points: Add to the previous model also change in the other to see whether my adaptation is influenced by the other's adaptation.

```
For colon we get conditionConversation:HR_self, conditionSynchronous:HR_self and conditionTurnTaking:HR_self significant
For star we only get HR_self and conditionSynchronous:HR_self significant. 

Interpreting:
We don't care about the intercepts. 
We care about stability as a sanity check. The heart rate should be stable. 
Probably significant p values. They are negative. cc:HRs. CT:HRs. CS:HRs
When we look at the other. It looks like not that much coupling effect. Does not seem to be significant coupling. 
If significant coupling: For synchronous it should be a positive number further from 1?? We have a standard error of only 0.01. 



Why are we predicting HR_change instead of HR_future?
The change will tell us whether it will e.g. keep going up. 

The fixed / structure of effect are stability, coupling and condition. 
We cannot expect coupling to be the same for all participants because one participant mght be much more engage. 
We cannot expect stability to be the same for all participants because for e.g. healthier people they have stronger HR. 

We hope there is an average statistical effect across all participants. There are good reasons for this --> therefore random effects. 

If very engage in a pair --> condition might change the value of coupling and stability which can differ across pairs. 

How do we model this?
1 ~ (HR_self + HR_other) * condition
This just gives one intercept. The average level of change when HR_self and HR_other are 0. 

We might want seperate intercepts for each group
We do this by adding 0 instead of 1. Meaning don't give me 1 intercept
0 ~ (HR_self + HR_other) * condition
When both are at baseline. What is the intercept for each condition. 3 values for us. 
0 + condition + (HR_s + HR_o):condition
Here we say 3 intercepts for main effect and 3 slopes(?) for the condition interaction.
The surrogate pairs are a better control baseline. Then we can add type.

0 + condition + (HR_s + HR_o):condition + 
   (0 + condition + (HR_self + HR_other):condition | participant) + #We tell the model not all data is the same. Counting for interdependence of the data
   (0 + condition + (HR_self + HR_other):condition | group) # We are doing partial pooling. Keep in mind there is another person in the pair. 
   ONLY IN THE IDEAL WORLD. Or when we learn bayesian statistics. Too complex. It's 1000 dimensions. In Bayes we can put in prior info. 
  
Instead we should simplyfy the model
We remove stability and coupling. 
0 + condition + (HR_s + HR_o):condition
   


## Now we need to create control baselines.

First shuffled controls, then surrogate pairs.

### Creating controls: shuffled controls

Shuffled controls break the temporal dependencies of time-series by shuffling the value within one time-series. This ensures the "coordination" observed is not due to the actual values in the series and not their sequence.
Tip: sample() is your friend, but make sure to shuffle things within participant/condition and not throughout the whole dataset
 
```{r}

# Create a shuffled dataset


# Concatenate it to the original dataset (and remember to have a column telling you which is which)
shuffle <- dd
shuffle$HR_self = sample(dd$HR_self)
shuffle$HR_other = sample(dd$HR_other)
shuffle$HR_self_past = lag(dd$HR_self, 1)
shuffle$HR_change_self = dd$HR_self_past-dd$HR_self
shuffle$pair <- "shuffle"
dd$pair <- "real"
dd <- merge(dd, shuffle, all = T)

# Create the same models as in the previous chunk, but adding an interaction by shuffled vs. real



# We use : to better interpret the interaction. We add parenthesis to see the effect compared to our previous. 
model_shuffle <- lmer(HR_change_self ~ 0 + condition + 
                ((HR_self_past + HR_other_past):condition):pair + 
                (0 + condition | participant) +
                (0 + condition | group), data = dd)

#Also run it with the star * so we can see whether the interaction is there

model_shuffle_s <- lmer(HR_change_self ~ 0 + condition + 
                ((HR_self_past + HR_other_past)*condition)*pair + 
                (0 + condition | participant) +
                (0 + condition | group), data = dd)

#I get no p values
summary(model_shuffle)
summary(model_shuffle_s)


```
 
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair)

```{r}
# Identify unique pairs within a given study (to keep things manageable) and create list of possible surrogate pairs (e.g. individual 1 from pair 1 and individual 2 from pair 2)

# Starting from the wide format, create "surrogate" dataset with the data from surrogate pairs
Groups <- as.numeric(as.character(unique(dd$group[dd$study==4]))) # List all pairs

SurrogateList <- expand.grid(a = Groups, b = Groups) # Identify all possible combinations of 2 pairs
SurrogateList = subset(SurrogateList, a != b)  # exclude combinations with identical pairs

for (i in 1:nrow(SurrogateList)){  # loop through all combinations
      x <- subset(dd, group==SurrogateList$a[i]) # subset data from the first pair
      y <- subset(dd, group==SurrogateList$a[i]) # subset data from the second pair    
      group <- c(800 + ((1:4)*i))                # create new pair id
      for (co in c("Synchronous","TurnTaking", "Conversation")){ # loop through conditions
         if (co %in% unique(x$condition) & co %in% unique(y$Condition)){ # check that both pairs have the data for that condition           
           z1 <- subset(x, condition==co)   # subset only that condtion from first pair         
           z2 <- subset(y, condition==co)   # subset only that condtion from second pair
           if (nrow(z1) > nrow(z2)) {	# make sure data have same length in both pairs
                z1<-z1[1:length(z2)]
              }          
          if (nrow(z2) > nrow(z1)) {z2<-z2[1:length(z1)]              
                 }
               w1 <- z1 %>% mutate(	# assemble new pair combining the 2 pairs                  
                 HR2 = z2$HR2, 
                 Resp2 = z2$Resp2, 
                 HR2_lead = z2$HR2_lead, 
                 Resp2_lead =z2$Resp2_lead, 
                 HR2_change = z2$HR2_change, 
                 Resp2_change = z2$Resp2_change)
                  write.csv(w1, "w1.csv")
               ### DO SOMETHING TO SAVE THIS	# make sure that you do this
             } }}
w1 <- read.csv("w1.csv")      
# Make it into long format

# Create models as in chunks above, but adding an interaction with the Real vs. Surrogate variable (exclude shuffled ones for simplicity)
#We add type ad an interaction
# 0 + (condition + (HR_s + HR_o):Condition):type
#For conversation the coupling parameter is probably different



```
 

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them

 